---
title: "Mapping urban water data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mapping-urban-water-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(uwdtap)
```

# Mapping county-level urban water data

The /code{"map_data"} function allows users to visualize differences in water data by county. Our hope is that by connecting publicly-accessible reports with mapping functionality, water data could be qualitatively connected with other geographically-linked environmental and social analyses. For example, geospatial data is used in analyses relating to [Social Determinants of Health](https://www.cdc.gov/places/social-determinants-of-health-and-places-data/index.html) and [water quality](https://www.epa.gov/waterdata/water-quality-data-download). Further, data gaps have been cited as a barrier in social and economic dimensions of water in California, suggesting that increasing accessibility to machine-readabale, transparent, and publicly available datasets, including those regarding water supply, could further such analyses (https://www.liebertpub.com/doi/full/10.1089/ees.2020.0315, Goddard, Ray, Balazs).  
